---
title: "Class Project"
author: "Mystery"
date: "November 10, 2024"
output: html_notebook
---
### Guiding Question (i.e., Research Question):
Are the dog breeds with longer lifespans more popular than dog breeds with shorter lifespans?

### Become acquainted with your data sources: 

### Where did you find them? 
I found the data on the websites for the American Kennel Club (AKC) and Spot Pet Insurance.

### Who collected/maintains them? 
The primary data source is collected from data on the AKC website and maintained in github by "tmfilho". The data on Spot's website is sourced from the AKC and Reader's Digest.

### When and why were they originally collected? 
The primary data source was collected to show information on 277 dog breeds including their lifespan and different traits. The data on most popular breeds from the AKC website was last updated in 2020 to show the most popular breeds by year. The data from Spot was last updated on 9/16/24 to show the most popular breed in each state.

### What does a case represent in each data source, and how many total cases are available? 
The primary source (AKC_Data) has 277 cases of individual dog breeds. I have two secondary sources. The first (PopularBreeds) has 192 cases. Each case is the ranking of and individual dog breed from 2014 to 2018. The second (StateBreeds) has 50 cases. Each case is the first and second most popular dog breed in an individual state in the US.

### What are some of the variables that you plan to use? 
From the primary source I will use variables for dog breed, life expectancy, and weight. From the secondary sources I will use breed, state, year, and popularity.

```{r}
rm(list = ls())
library(rvest)
library(dplyr)
library(lubridate)
library(readr)
library(tidyverse)
library(readr)
```


```{r}
AKC_Data<- read_csv("akc-data-2020-05-17 17_35_49.397041.csv") ###downloaded from https://tmfilho.github.io/akcdata/

AKC_Data <- AKC_Data%>%
select(...1, temperament, popularity, min_height, max_height, min_weight, max_weight, min_expectancy, max_expectancy, group, grooming_frequency_value, grooming_frequency_category, shedding_value, shedding_category, energy_level_value, energy_level_category, trainability_value, trainability_category, demeanor_value, demeanor_category)%>%
rename("Breed" = ...1)

str(AKC_Data)
head(AKC_Data)
```

```{r}
page <- "https://www.akc.org/expert-advice/news/most-popular-dog-breeds-full-ranking-list/"
BreedPop <- page %>%
  read_html() %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)

PopularBreeds <- BreedPop[[1]]
str(PopularBreeds)
head(PopularBreeds)
```

```{r}
pagetwo <- "https://spotpet.com/blog/breed-tips/most-popular-dog-breeds-by-state"
ByState <- pagetwo %>%
  read_html() %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)

StateBreeds <- ByState[[1]]
str(StateBreeds)
head(StateBreeds)
```

```{r}
StateBreeds <- StateBreeds%>%
  filter(row_number() >1)%>%
  rename("State" = X1, "1st Most Popular" = X2, "2nd Most Popular" = X3)

StateBreeds  
```

### Explore intuition related to the research question:

### Create some informative plots and summary statistics.

Below we will calculate the average weight and average life expectancy from the minimums and maximum values recorded for each category.

```{r}
Lifespan <- AKC_Data%>%
  select(Breed, min_expectancy, max_expectancy, min_weight, max_weight)%>%
  mutate(AVG_Lifespan = (min_expectancy + max_expectancy)/2)%>%
  mutate(AVG_Weight = (min_weight + max_weight)/2)%>%
  arrange(Breed)

Lifespan
```

### In this graph we can see that the larger breeds tend to have shorter lifespans than the smaller breeds.

```{r warning=FALSE}
Lifespan%>%
  ggplot(aes(x=AVG_Weight, y=AVG_Lifespan))+
  stat_smooth()+
  geom_point()+
  theme_bw()
```


```{r warning=FALSE}
PopularBreeds <- PopularBreeds%>%
  mutate(`2018 Rank` = as.numeric(`2018 Rank`))%>%
  mutate(`2017 Rank` = as.numeric(`2017 Rank`))%>%
  mutate(`2016 Rank` = as.numeric(`2016 Rank`))%>%
  mutate(`2015 Rank` = as.numeric(`2015 Rank`))%>%
  mutate(`2014 Rank` = as.numeric(`2014 Rank`))
```

### After ensuring columns are changed from character strings to numeric, I can then find the average popularity of each dog breed


```{r}
Avg_Pop <- PopularBreeds%>%
  mutate(AVG_Rank = rowMeans(cbind(`2018 Rank`, `2017 Rank`, `2016 Rank`, `2015 Rank`, `2014 Rank`), na.rm = TRUE))%>%
  arrange(AVG_Rank)

Avg_Pop
```

### Describe preliminary observations and intuition about the research question.

From the wrangling I have completed with the data sets so far I can see that smaller breeds tend to live longer, but the three most popular dog breeds are large dogs. From this preliminary information it would seem that breeds that live longer are not more popular with pet owners than breeds with shorter lifespans. Further work with these data sets and looking at the bigger picture of all the data may show something different.

### I want to include a new column to list the average lifespan of the breed alongside the popularity for the given years. Putting the Avg_Pop table in alphabetical order makes it easier to check the two tables for compatablity. I can then see that the breed names in the Avg_Pop table are plural, while they are singular in the Lifespan table. To join them I will first have to clean up the breed names so they match properly

```{r}
Avg_Pop_Alpha <-
  Avg_Pop%>%
  arrange(Breed)

Avg_Pop_Alpha        ### Need to remove "s" from end of breed names to join with the AvgLifespan table
```

```{r}
Avg_Pop <- 
  Avg_Pop%>%
  mutate(Breed = gsub(pattern = "s$", replacement = "\\1", Breed))

Avg_Pop
```

```{r}
AvgLifespan <-
  Lifespan%>%
  select(Breed, AVG_Lifespan)

Avg_Life_Rank <-
  Avg_Pop%>%
  inner_join(AvgLifespan, by = c(Breed = "Breed"))

Avg_Life_Rank <-
  Avg_Life_Rank%>%
  select(Breed, AVG_Rank, AVG_Lifespan)

Avg_Life_Rank
```



### Below I have combined the breed average rank with the most popular breeds by state to see what additional information we can gain on the popularity of longer versus shorter lived breeds.

```{r}
Pop_By_State <-
  StateBreeds%>%
  select(State, `1st Most Popular`)

Pop_By_State

Pop_By_State <- Pop_By_State%>%
  mutate(`1st Most Popular` = gsub(pattern = "German Shepherd", replacement = "German Shepherd Dog", `1st Most Popular`)) ### adjusting breed name to join tables
  
State_Rank_Life <-
  Pop_By_State%>%
  inner_join(Avg_Life_Rank, by = c(`1st Most Popular` = "Breed")) ### 2 states removed from list due to no breed match

State_Rank_Life

State_Rank_Life%>%
  arrange(desc(AVG_Lifespan))
```

### Now that we have the combined information on the most popular breed for each state as well as the average overall lifespan and popularity of these breeds we can see that the majority of the most popular breeds tend to have shorter lifespans. 
```{r}
ggplot(State_Rank_Life) +
  aes(x = AVG_Rank, y = AVG_Lifespan) +
  geom_point() +
  theme_minimal() +
  facet_wrap(vars(`1st Most Popular`))

```
  
```{r}
Lifespan%>%
  summarise(median(AVG_Lifespan, na.rm = TRUE))

Size_Life <- Lifespan%>%
  mutate(Size = if_else(AVG_Weight <30, "small", "large", missing = NULL))%>%
  mutate(Lived = if_else(AVG_Lifespan <12, "short", "long", missing = NULL))

Size_Life <- Size_Life%>%
  select(Breed, Size, Lived)
Size_Life
```

```{r}
All_Info <- State_Rank_Life%>%
  inner_join(Size_Life, by = c(`1st Most Popular` = "Breed"))%>%
  rename(Breed = `1st Most Popular`)

All_Info
```
### Finally, we can see below that only two breeds that are considered "long-lived" have a ranking in the top 10. They are the Beagle and the Yorkshire Terrier. The Labrador Retriever is overwhelmingly the most popular breed reported in our data, but we can also see high popularity for the German Shepherd Dog and the Golden Retriever. All 3 of these breeds would be considered large dogs with relatively short lifespans. Further questions we could consider pursuing with more data would be looking at other factors like whether the dogs are popular as house pets or working dogs. The prevelance of dogs used in jobs such as herding and guarding livestock, hunting, and work as service or assistance animals could skew popularity. If more information on the types of households these dogs inhabit was available we could dive further into defining the most popular pet breeds versus the most popular working breeds.

```{r}
ggplot(All_Info) +
  aes(x = AVG_Rank, color = Breed, fill = Breed) +
  geom_histogram(bins = 30L) +
  theme_bw() +
  facet_wrap(vars(Lived))
```
