---
title: "Class Project"
author: "Elizabeth Shaw"
output: html_notebook
---
### Guiding Question (i.e., Research Question):
Make a clear statement of a single guiding/research question that will drive your analysis. It should be phrased as a question that you will try to investigate (i.e., it should have a question mark). 

### Become acquainted with your data sources: 
Where did you find them? 
Who collected/maintains them? 
When and why were they originally collected? 
What does a case represent in each data source, and how many total cases are available? 
What are some of the variables that you plan to use? 

```{r}
rm(list = ls())
library(rvest)
library(dplyr)
library(lubridate)
library(readr)
library(tidyverse)
library(readr)
```


```{r}
AKC_Data<- read_csv("akc-data-2020-05-17 17_35_49.397041.csv")

AKC_Data <- AKC_Data%>%
select(...1, temperament, popularity, min_height, max_height, min_weight, max_weight, min_expectancy, max_expectancy, group, grooming_frequency_value, grooming_frequency_category, shedding_value, shedding_category, energy_level_value, energy_level_category, trainability_value, trainability_category, demeanor_value, demeanor_category)%>%
rename("Breed" = ...1)

str(AKC_Data)
head(AKC_Data)
```

```{r}
page <- "https://www.akc.org/expert-advice/news/most-popular-dog-breeds-full-ranking-list/"
BreedPop <- page %>%
  read_html() %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)

PopularBreeds <- BreedPop[[1]]
str(PopularBreeds)
head(PopularBreeds)
```

```{r}
pagetwo <- "https://spotpet.com/blog/breed-tips/most-popular-dog-breeds-by-state"
ByState <- pagetwo %>%
  read_html() %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)

StateBreeds <- ByState[[1]]
str(StateBreeds)
head(StateBreeds)
```

```{r}
StateBreeds <- StateBreeds%>%
  filter(row_number() >1)%>%
  rename("State" = X1, "1st Most Popular" = X2, "2nd Most Popular" = X3)

StateBreeds  
```

