---
title: "Class Project"
author: "Elizabeth Shaw"
date: "November 10, 2024"
output: html_notebook
---
### Guiding Question (i.e., Research Question):
Are the dog breeds with longer lifespans more popular than dog breeds with shorter lifespans?

### Become acquainted with your data sources: 

### Where did you find them? 
I found the data on the websites for the American Kennel Club (AKC) and Spot Pet Insurance.

### Who collected/maintains them? 
The primary data source is collected from data on the AKC website and maintained in github by "tmfilho". The data on Spot's website is sourced from the AKC and Reader's Digest.

### When and why were they originally collected? 
The primary data source was collected to show information on 277 dog breeds including their lifespan and different traits. The data on most popular breeds from the AKC website was last updated in 2020 to show the most popular breeds by year. The data from Spot was last updated on 9/16/24 to show the most popular breed in each state.

### What does a case represent in each data source, and how many total cases are available? 
The primary source (AKC_Data) has 277 cases of individual dog breeds. I have two secondary sources. The first (PopularBreeds) has 192 cases. Each case is the ranking of and individual dog breed from 2014 to 2018. The second (StateBreeds) has 50 cases. Each case is the first and second most popular dog breed in an individual state in the US.

### What are some of the variables that you plan to use? 
From the primary source I will use variables for dog breed, life expectancy, and weight. From the secondary sources I will use breed, state, year, and popularity.

```{r}
rm(list = ls())
library(rvest)
library(dplyr)
library(lubridate)
library(readr)
library(tidyverse)
library(readr)
```


```{r}
AKC_Data<- read_csv("akc-data-2020-05-17 17_35_49.397041.csv")

AKC_Data <- AKC_Data%>%
select(...1, temperament, popularity, min_height, max_height, min_weight, max_weight, min_expectancy, max_expectancy, group, grooming_frequency_value, grooming_frequency_category, shedding_value, shedding_category, energy_level_value, energy_level_category, trainability_value, trainability_category, demeanor_value, demeanor_category)%>%
rename("Breed" = ...1)

str(AKC_Data)
head(AKC_Data)
```

```{r}
page <- "https://www.akc.org/expert-advice/news/most-popular-dog-breeds-full-ranking-list/"
BreedPop <- page %>%
  read_html() %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)

PopularBreeds <- BreedPop[[1]]
str(PopularBreeds)
head(PopularBreeds)
```

```{r}
pagetwo <- "https://spotpet.com/blog/breed-tips/most-popular-dog-breeds-by-state"
ByState <- pagetwo %>%
  read_html() %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)

StateBreeds <- ByState[[1]]
str(StateBreeds)
head(StateBreeds)
```

```{r}
StateBreeds <- StateBreeds%>%
  filter(row_number() >1)%>%
  rename("State" = X1, "1st Most Popular" = X2, "2nd Most Popular" = X3)

StateBreeds  
```

### Explore intuition related to the research question:

### Create some informative plots and summary statistics.

```{r}
Lifespan <- AKC_Data%>%
  select(Breed, min_expectancy, max_expectancy, min_weight, max_weight)%>%
  mutate(AVG_Lifespan = (min_expectancy + max_expectancy)/2)%>%
  mutate(AVG_Weight = (min_weight + max_weight)/2)%>%
  arrange(Breed)

Lifespan
```

```{r}
Lifespan%>%
  ggplot(aes(x=AVG_Weight, y=AVG_Lifespan))+
  geom_point()+
  theme_bw()
```


```{r}
PopularBreeds <- PopularBreeds%>%
  mutate(`2018 Rank` = as.numeric(`2018 Rank`))%>%
  mutate(`2017 Rank` = as.numeric(`2017 Rank`))%>%
  mutate(`2016 Rank` = as.numeric(`2016 Rank`))%>%
  mutate(`2015 Rank` = as.numeric(`2015 Rank`))%>%
  mutate(`2014 Rank` = as.numeric(`2014 Rank`))
```




```{r}
Avg_Pop <- PopularBreeds%>%
  mutate(AVG_Rank = rowMeans(cbind(`2018 Rank`, `2017 Rank`, `2016 Rank`, `2015 Rank`, `2014 Rank`), na.rm = TRUE))%>%
  arrange(AVG_Rank)

Avg_Pop
```

### Describe preliminary observations and intuition about the research question.

From the wrangling I have completed with the data sets so far I can see that smaller breeds tend to live longer, but the three most popular dog breeds are large dogs. From this preliminary information it would seem that breeds that live longer are not more popular with pet owners than breeds with shorter lifespans. Further work with these data sets and looking at the bigger picture of all the data may show a different picture when looking at all data together.


